Effects of the body force on the pedestrian and evacuation dynamics

### 2020-02-06 paper.tex
reemplace ~ por \,

We confirm from these diagrams that no univocal relations can be established 
between the individual parameters and the collective dynamics (in a crowded 
environment). The presence of the body force moves the dynamics to a more 
complex context. We will investigate this context in section~\ref{results}. \\

@@

Despite the presence of the FIS or FIF pattern, the evacuation time at a 
fixed value of $v_d$ decreases for increasing values of $k_n$ (whiting the examined interval). 
This means that stiffer pedestrians evacuate faster than soft pedestrians. 
To further investigate this 
phenomenon, we computed the mean velocity of the whole crowd as a function of the 
stiffness $k_n$. Fig.~\ref{kn_vs_vx_bottleneck} shows the actual mean velocity in the 
$x-$direction as a function of the stiffness $k_n$ for three different desired 
velocities. In all cases, the velocity increases with the stiffness. The most significative 
increment of velocity is in the interval $k_n>$E4. \\

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_bottleneck.eps}}\\
\caption[width=0.47\columnwidth]{Mean velocity in the $x-$direction as a function 
of the stiffness level $k_n$ for different desired velocities (see label). The 
data corresponds to a bottleneck with periodic boundary conditions (re-injecting 
pedestrians). The average was taken every five seconds once the crowd reaches the 
stationary state ($t=$20~s) until the end of the simulation ($t=$1000~s). 
Color online only. }
\label{kn_vs_vx_bottleneck}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Snapshots of the contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)$~m, the width of the door is 0.92~m (equivalent to 2 pedestrian\textsc{\char13}s diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). (a) and (b) correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. (c) and (d) correspond to simulations with $k_n$=1.2\,E6 with $v_d=$2 and $v_d=$10 respectively. Color online only.}
\label{network_bottleneck}
\end{figure}

The four configurations corresponds to two different $v_d$ and two different $k_n$ values (say, the  minimum and maximum explored values). Fig.~\ref{network_vd2_kn0} and Fig.~\ref{network_vd10_kn0} show snapshots for $k_n=0$, at the desired velocities of 2$\,$m/s and 10$\,$m/s, respectively.    Fig.~\ref{network_vd2_kn1200000} and Fig.~\ref{network_vd10_kn1200000} show similar situations, but for  $k_n=$1.2$\,$E6. As expected, increasing the desired velocity compresses the crowd towards the exit. \\

@@

In order to get a better insight on how the pedestrians contact to each other, we present in Fig.~\ref{network_corridor} two snapshots of the corridor at the stationary situation. Fig.~\ref{network_d6_knE5} corresponds to $k_n=1.2$~E5. It can be seen that increasing the stiffness increases the number of contacts between pedestrians (say, the network connectivity). The number of nodes attaining degree 6 is far larger in the case of $k_n=1.2\,$E5 than the case of $k_n=0$. \\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2$~E5. The friction coefficient and the other SFM parameters are the same as in Section \ref{bottleneck}. Color online only.}
\label{network_corridor}
\end{figure}

@@

The most important result is that increasing the $k_n$ value reduces the flow (hence reduces the walking speed of the overall crowd) in the corridor geometry. Fig.~\ref{kn_vs_vx_corridor} shows the mean velocity in the direction of displacement $v_x$ as a function of the stiffness for different global density levels. $v_x$ reduces as $k_n$ increases as opposed to what happens in the bottleneck geometry. There is little to no change in the speed from zero to $k_n=$2.6~E4, from this value on, the diminishing of velocity is more abrupt.\\

The reason why the velocity reduces as the stiffness of the pedestrians' increases is that it becomes harder for pedestrians to walk forward. The difficulty arises because the pedestrian feels a much higher repulsion from the pedestrian ahead (due to the body force increment). Thus, they have no choice but to move at (almost) the same speed as the individuals they have at their sides. This microscopic mechanism explains the macroscopic behavior that turns the crowd from a ``fluid-like state'' into a ``solid-like state''. When the stiffness is high enough (say $k_n=$1.2~E6), all the pedestrians walk at the same velocity. The pedestrians that walk in physical contact with the wall are the ones who determine the velocity of the whole crowd.\\

@@

The value of $k_n$ affects the velocity all along the corridor. Fig.~\ref{vx_profile} shows the velocity profile ($v_x$ vs. the longitudinal coordinate $y$) for different $k_n$ values. For $k_n=0$, we can see a parabolic-like velocity profile that resembles the Poiseuille flow (for Newtonian and incompressible fluids in laminar flow flowing through a pipe). This behavior was observed in empirical measurements reported in Ref.~\cite{zhang2013empirical}. As $k_n$ increases the velocity profile flattens until becoming a constant ($k_n=$ 1.2~E6). In this scenario, the value of mean $v_x$ is much lower than on the case of soft pedestrians ($k_n=$0).\\

@@

Figs. \ref{flow_density} show the flow as a function of the global density (fundamental diagram). Fig.~\ref{flow_density_no_body_force} corresponds to $\mathcal{K}_c=$ 0 (zero body force \textit{i.e.} $k_n=0$) while Fig.~\ref{flow_density_body_force} corresponds to $\mathcal{K}_c=$68 (which is the value of the original SFM with $k_n$=1.2~E5 ). Each curve correspond to a different friction value, hence different $k_t$ value (see the caption).\\

@@

\begin{comment}
Fig. \ref{flow_density} shows the flow as a function of the global density (fundamental diagram). \ref{flow_density_no_body_force} corresponds to zero body force (\textit{i.e.} $k_n=0$), Fig.~\ref{flow_density_body_force} corresponds to a body force with $k_n$=1.2~E5. Each curve correspond to a different friction value (see the caption).\\

>>>>

We confirm from these diagrams that no univocal relations can be established 
between the individual parameters and the collective dynamics (in a crowded 
environment). The presence of the body force moves the dynamics to a more 
complex context. We will investigate this context in Section~\ref{results}. \\

@@

Despite the presence of the FIS or FIF pattern, the evacuation time at a 
fixed value of $v_d$ decreases for increasing values of $k_n$ (whiting the examined interval). 
This means that stiffer pedestrians evacuate faster than soft pedestrians. 
To further investigate this 
phenomenon, we computed the mean velocity of the whole crowd as a function of the 
stiffness $k_n$. Fig.~\ref{kn_vs_vx_bottleneck} shows the actual mean velocity in the 
$x-$direction as a function of the stiffness $k_n$ for three different desired 
velocities. In all cases, the velocity increases with the stiffness. The most significative 
increment of velocity is in the interval $k_n>\,$E4. \\

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_bottleneck.eps}}\\
\caption[width=0.47\columnwidth]{Mean velocity in the $x-$direction as a function 
of the stiffness level $k_n$ for different desired velocities (see label). The 
data corresponds to a bottleneck with periodic boundary conditions (re-injecting 
pedestrians). The average was taken every five seconds once the crowd reaches the 
stationary state ($t=20\,$s) until the end of the simulation ($t=1000\,$s). 
Color online only. }
\label{kn_vs_vx_bottleneck}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Snapshots of the contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)$~m, the width of the door is 0.92~m (equivalent to 2 pedestrian\textsc{\char13}s diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). (a) and (b) correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. (c) and (d) correspond to simulations with $k_n=1.2\,$E6 with $v_d=$2 and $v_d=$10 respectively. Color online only.}
\label{network_bottleneck}
\end{figure}

The four configurations corresponds to two different $v_d$ and two different $k_n$ values (say, the  minimum and maximum explored values). Fig.~\ref{network_vd2_kn0} and Fig.~\ref{network_vd10_kn0} show snapshots for $k_n=0$, at the desired velocities of 2$\,$m/s and 10$\,$m/s, respectively.    Fig.~\ref{network_vd2_kn1200000} and Fig.~\ref{network_vd10_kn1200000} show similar situations, but for  $k_n=1.2\,$E6. As expected, increasing the desired velocity compresses the crowd towards the exit. \\

@@

In order to get a better insight on how the pedestrians contact to each other, we present in Fig.~\ref{network_corridor} two snapshots of the corridor at the stationary situation. Fig.~\ref{network_d6_knE5} corresponds to $k_n=1.2\,$E5. It can be seen that increasing the stiffness increases the number of contacts between pedestrians (say, the network connectivity). The number of nodes attaining degree 6 is far larger in the case of $k_n=1.2\,$E5 than the case of $k_n=0$. \\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2\,$E5. The friction coefficient and the other SFM parameters are the same as in Section \ref{bottleneck}. Color online only.}
\label{network_corridor}
\end{figure}

@@

The most important result is that increasing the $k_n$ value reduces the flow (hence reduces the walking speed of the overall crowd) in the corridor geometry. Fig.~\ref{kn_vs_vx_corridor} shows the mean velocity in the direction of displacement $v_x$ as a function of the stiffness for different global density levels. $v_x$ reduces as $k_n$ increases as opposed to what happens in the bottleneck geometry. There is little to no change in the speed from zero to $k_n=2.6\,$E4, from this value on, the diminishing of velocity is more abrupt.\\

The reason why the velocity reduces as the stiffness of the pedestrians' increases is that it becomes harder for pedestrians to walk forward. The difficulty arises because the pedestrian feels a much higher repulsion from the pedestrian ahead (due to the body force increment). Thus, they have no choice but to move at (almost) the same speed as the individuals they have at their sides. This microscopic mechanism explains the macroscopic behavior that turns the crowd from a ``fluid-like state'' into a ``solid-like state''. When the stiffness is high enough (say $k_n=1.2\,$E6), all the pedestrians walk at the same velocity. The pedestrians that walk in physical contact with the wall are the ones who determine the velocity of the whole crowd.\\

@@

The value of $k_n$ affects the velocity all along the corridor. Fig.~\ref{vx_profile} shows the velocity profile ($v_x$ vs. the longitudinal coordinate $y$) for different $k_n$ values. For $k_n=0$, we can see a parabolic-like velocity profile that resembles the Poiseuille flow (for Newtonian and incompressible fluids in laminar flow flowing through a pipe). This behavior was observed in empirical measurements reported in Ref.~\cite{zhang2013empirical}. As $k_n$ increases the velocity profile flattens until becoming a constant ($k_n=1.2\,$E6). In this scenario, the value of mean $v_x$ is much lower than on the case of soft pedestrians ($k_n=$0).\\

@@

Figs. \ref{flow_density} show the flow as a function of the global density (fundamental diagram). Fig.~\ref{flow_density_no_body_force} corresponds to $\mathcal{K}_c=$ 0 (zero body force \textit{i.e.} $k_n=0$) while Fig.~\ref{flow_density_body_force} corresponds to $\mathcal{K}_c=$68 (which is the value of the original SFM with $k_n=1.2\,$E5 ). Each curve correspond to a different friction value, hence different $k_t$ value (see the caption).\\

@@

\begin{comment}
Fig. \ref{flow_density} shows the flow as a function of the global density (fundamental diagram). \ref{flow_density_no_body_force} corresponds to zero body force (\textit{i.e.} $k_n=0$), Fig.~\ref{flow_density_body_force} corresponds to a body force with $k_n=1.2\,$E5. Each curve correspond to a different friction value (see the caption).\\
